---
title: "R Basics"
output: html_notebook
---

## Your Turn 1

Create a character vector of 3 fruit names and assign it to the variable `fruits`. On the next line, type the variable name to view its contents.

```{r}

```

## Your Turn 2

Bring up the help file for the `seq()` function. Then, use the `seq()` function to generate a numeric vector from 1 to 8 by 0.5. (1, 1.5 ... 7.5, 8) (the _examples_ section at the bottom is probably the most useful). Assign it to a variable called `numbers`. On the next line, type the variable name to view its contents.

```{r}

```

## Your Turn 3

How many elements are in the vectors produced by:

- `max(numbers)`
- `min(numbers)`
- `mean(numbers)`
- `median(numbers)`
- `sum(numbers)`
- `numbers * 3`
- `numbers / numbers`
- `numbers - 2`
- `str_c("Fruit", fruits, sep = ": ")`
- `numbers >= 4`
- `numbers != 4`
- `fruits == "banana"`
- `fruits %in% c("banana", "pear")`

```{r}

```

## Your Turn 4

The following code returns `NA`. Fix it to get the mean of all the non-missing arguments.

```{r}
mean(c(1, 2, 3, NA))
```

***

# Take Aways

- Use the `<-` operator to assign the result of an expression to a variable
- Functions have arguments that can be specified by position or by name, and are separated by a comma `,`.
- Many functions in R operate on vectors, and return a vector of the same length ("vectorized"), or length 1 ("summary").
- `NA` or missing values are propogated through calculations unless you specifically pass `na.rm = TRUE`.
- You can bring up the help file for a function using `?function_name` or `help(function_name)`

# More Practice

- The [Workflow: Basics](https://r4ds.had.co.nz/workflow-basics.html) chapter of [R for Data Science](https://r4ds.had.co.nz/workflow-basics.html).
- The [Programming Basics](https://rstudio.cloud/learn/primers/1.2) primer from [RStudio Cloud](https://rstudio.cloud/).
