---
title: "Reading Data in R (solutions)"
output: html_notebook
---

```{r setup}
library(tidyverse)
```

## Your Turn 1

Reading an Excel file requires the `read_excel()` function from the _readxl_ package. Load the _readxl_ package below (it is already installed on your computer if you installed the _tidyverse_ package):

```{r}

```

## Your Turn 2

Use the `read_excel()` function to read the `"warwick.xlsx"` file, and assign it to a variable called `warwick_xl`.

```{r}

```

## Your Turn 3

The following code will read the `"warwick.csv"` file and assign it to the variable `warwick_csv`. Use the output from the command to set the `col_types` argument of `read_csv()`. Did it guess the column types correctly?

```{r}
warwick_csv <- read_csv(
  "warwick.csv"
)
```

## Your Turn 4

Use the `View()` function (i.e, click on the object name in the _Environment_ tab of RStudio), the `head()` function, and the `glimpse()` function to examine the three `warwick*` objects. Are they identical to each other (hint: there is only one difference)?

```{r}

```

***

# Take Aways

- The `read_excel()` function from the _readxl_ package can be used to read excel files as R data frames.
- The `read_csv()` function can be used to import CSV files into R as data frames. Use `col_types = cols(...)` to specify which columns are of which type.
- The `View()`, `head()` and `glimpse()` functions can be used to make sure the data imported properly.

# More Practice

- The [Data Import](https://r4ds.had.co.nz/data-import.html) chapter of [R for Data Science](https://r4ds.had.co.nz/)
- The [readxl package](https://readxl.tidyverse.org/) tutorials.
