---
title: "Visualize Data (solutions)"
output: html_document
---

```{r setup}
library(tidyverse)
library(geoscidata)
```

```{r}
warwick
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
# remove
ggplot(data = warwick) +
  geom_point(mapping = aes(x = Rb_ppm, y = K_ppm)) 
```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = warwick) +
  geom_point(mapping = aes(x = Rb_ppm, y = K_ppm, color = rock_type))
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(warwick) + geom_boxplot(aes(rock_name, K_ppm))
```

## Your Turn 4

Make a histogram of the `K_ppm` variable from `warwick`.

```{r}
ggplot(warwick) + geom_histogram(aes(K_ppm))
```

## Your Turn 5

Make a density plot of `K_ppm` colored by `rock_type`.

```{r}
ggplot(warwick) + geom_density(aes(K_ppm, color = rock_type))
```

## Your Turn 6

Make a bar chart of `rock_type` colored by `rock_type`.

```{r}
ggplot(warwick) + geom_bar(aes(x = rock_type, fill = rock_type))
```

## Your Turn 7

Predict what this code will do. Then run it.

```{r}
ggplot(warwick) + 
  geom_point(aes(Rb_ppm, K_ppm)) +
  geom_smooth(aes(Rb_ppm, K_ppm))
```

## Your Turn 8

What does `getwd()` return?

```{r}
getwd()
```

## Your Turn 9

Save the last plot and then locate it in the files pane.

```{r}
ggsave("my_plot.pdf", height = 6, width = 6)
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

# More practice

- The [Data Visualization](https://r4ds.had.co.nz/data-visualisation.html) chapter of [R for Data Science](https://r4ds.had.co.nz/)
- The [Data Visualization Primer](https://rstudio.cloud/learn/primers/3) from [RStudio Cloud](http://rstudio.cloud)
