---
title: "Visualizing Data (round 2) (solutions)"
output: html_document
---

```{r setup}
library(tidyverse)
library(geoscidata)
```

```{r}
warwick
```

## Your Turn 1

Modify the following code to use one of ggplot2's built-in themes with a base font size of 10 points. Experiment! See `?theme_bw` for a list of themes that can be used.

```{r}
ggplot(warwick) + 
  geom_point(aes(x = Fe_ppm, y = Mn_ppm, color = rock_type)) +
  # could also be theme_gray(), theme_dark(), theme_light(), theme_minimal(), theme_linedraw(), theme_classic()
  theme_bw(10)
```

You can add `theme()` modifications after you've set the global theme with one of the above themes.

## Your Turn 2

Set the `palette` argument of `scale_color_brewer()` to one of the values listed in the "Palettes" section of the help file (`?scale_color_brewer`). You can also preview them at http://colorbrewer2.org/ .

```{r}
?scale_color_brewer
ggplot(warwick) + 
  geom_point(aes(x = Fe_ppm, y = Mn_ppm, color = rock_type)) +
  scale_color_brewer(palette = "Spectral")
```

- Discrete scales also have a `limits` argument (which is one way to control the order of the values), and `labels` argument, which controls the display values. Both take a vector created with `c()`.

## Your Turn 3

Change the X and Y scales to log base 10 scales using the `scale_*_log10()` functions.

```{r}
ggplot(warwick) + 
  geom_point(aes(x = Fe_ppm, y = Mn_ppm, color = rock_type)) +
  scale_x_log10() +
  scale_y_log10()
```

- Continuous scales have a `breaks` argument, which specifies the location of the breaks. The `labels` argument specifies the values that are displayed next to the breaks. Both are specified as vectors using the `c()` function.

## Your Turn 4

Use the `facet_wrap()` function to produce a plot with four panels, one for each `rock_type`. Experiment with `scales = "free_x"`, `scales = "free_y"`, and `scales = "free"`.

```{r}
ggplot(warwick) + 
  geom_point(aes(x = Fe_ppm, y = Mn_ppm, color = rock_type)) +
  facet_wrap(~rock_type, scales = "free")
```

## Your Turn 5

Modify the `xlim` argument of `coord_cartesian()` to "zoom in" to the exclude the two outlying points with high Fe values.

```{r}
ggplot(warwick) + 
  geom_point(aes(x = Fe_ppm, y = Mn_ppm, color = rock_type)) +
  coord_cartesian(xlim = c(0, 250000))
```

## Exam

Create a scatterplot of `warwick` using the `Ti_ppm` and `Mn_ppm` variables, colored by `rock_name`. Facet this plot by the `rock_type` variable, and apply your favourite built-in theme. (For bonus points, include log x and log y scales).

```{r}
ggplot(warwick) +
  geom_point(aes(x = Ti_ppm, y = Mn_ppm, color = rock_name)) +
  facet_wrap(~rock_type) +
  theme_bw() +
  labs(x = "Ti (ppm)", y = "Mn (ppm)", color = "Rock Name") +
  scale_x_log10() + scale_y_log10()
```

***

# Take aways

- Themes control the non-data appearance of visual elements. Built-in themes can be added using `theme_*()` functions, and change the appearance of the whole plot.
- The default scales can be modified using `scale_<AESTHETIC>_*()` functions. This includes the x-axis, y-axis, and other aesthetic scales.
- Using `facet_wrap(~<COLUMN>)` is one way to split data into groups according to each value in`<COLUMN>`. The `scales` argument changes whether x scales, y scales, or both are identical in each panel.
- Zoom in to a particular part of the plot by setting the `xlim` and `ylim` arguments of `coord_cartesian()`, or flip the x and y axes using `coord_flip()`.
- The `labs(<AESTHETIC> = "<TEXT>")` function sets the text that labels the scale for each aesthetic.

# More Practice

- The [Exploratory Data Analysis](https://r4ds.had.co.nz/exploratory-data-analysis.html) and [Graphics for Communication](https://r4ds.had.co.nz/graphics-for-communication.html) chapters in [R for Data Science](https://r4ds.had.co.nz/).
- The [Exploratory Data Analysis Primer](https://rstudio.cloud/learn/primers/3.1) and the [Plot Customization Primer](https://rstudio.cloud/learn/primers/3.8) from [RStudio Cloud](https://rstudio.cloud/) from [RStudio Cloud](https://rstudio.cloud/).
