---
title: "R Notebook"
output: html_document
---

```{r setup}
library(tidyverse)
library(geoscidata)
```


## Your Turn 1

Create a scatterplot of `Mn_ppm` vs. `Fe_ppm`, with pretty labels for the `x` (`"Fe (ppm)"`), `y` (`"Mn (ppm)"`), and `color` (`"Rock Name"`) scales.

```{r}
ggplot(warwick) + 
  geom_point(aes(x = Fe_ppm, y = Mn_ppm, color = rock_name)) +
  labs(x = "Fe (ppm)", y = "Mn (ppm)", color = "Rock Name")
```

In addition to aesthetics like `x`, `y`, and `col`, you can use `title`, `subtitle`, and `caption` to include these elements in the plot portion of the output.

## Your Turn 2

Run the following two chunks to generate the two plots. What is the difference between them?

```{r}
ggplot(warwick) + 
  geom_point(aes(Rb_ppm, K_ppm), alpha = 0.1, color = "blue")
```

```{r}
ggplot(warwick) + 
  geom_point(aes(Rb_ppm, K_ppm, alpha = 0.1, color = "blue"))
```

## Your Turn 3

Predict what this code will do. Then run it.

```{r}
ggplot(warwick) + 
  geom_point(aes(Rb_ppm, K_ppm)) +
  facet_wrap(~rock_type)
```

## Your Turn 4

Modify the following code to use one of ggplot2's built-in themes with a base font size of 10 points.

```{r}
ggplot(warwick) +
  geom_point(aes(x = K_ppm, y = Rb_ppm), alpha = 0.1) +
  facet_wrap(~rock_type) +
  # could also be theme_bw(), theme_dark(), theme_light(), theme_minimal(), theme_linedraw()
  theme_bw(10)
```

There are more themes in the _ggthemes_ package (try some!).

***

# Take aways

- The `labs(<AESTHETIC> = "<TEXT>")` function sets the text that labels the scale for each aesthetic.
- Use `<GEOM_FUNCTION>(aes(<MAPPED_AESTETIC> = <COLUMN>), <SET_AESTHETIC> = <VALUE>)` to __set__ values of aesthetics that are the same for the entire layer. 
- `theme_*()` functions change the default appearance of graphic elements that are not related to the data.
- Using `facet_wrap(~<COLUMN>)` is one way to split data into groups according to each value in`<COLUMN>`

# More Practice

- The [Exploratory Data Analysis](https://r4ds.had.co.nz/exploratory-data-analysis.html) chapter in [R for Data Science](https://r4ds.had.co.nz/).
- The [Exploratory Data Analysis Primer](https://rstudio.cloud/learn/primers/3.1) from [RStudio Cloud](https://rstudio.cloud/)
