---
title: "Geology-specific plots (solutions)"
output: html_document
---

```{r setup}
library(tidyverse)
library(geoscidata)
```

## Basic Maps

The following code will create a map of the sample locations using the `ggspatial` package. Use what you know about ggplots to make the following modifications:

- Facet the plot by the `rock_type` variable using `facet_wrap()`
- Use the `labs()` function to give appropriate labels to the scales

```{r}
library(ggspatial)

ggplot(warwick) +
  annotation_map_tile() +
  geom_spatial_point(aes(longitude, latitude, color = rock_group_abbrev), alpha = 0.3) +
  annotation_scale() +
  facet_wrap(~rock_type) +
  labs(x = NULL, y = NULL, color = "Group") +
  theme(legend.position = "bottom")
```

For more on how to use the [ggspatial package](https://github.com/paleolimbot/ggspatial), see the [vignette on CRAN](https://cran.r-project.org/web/packages/ggspatial/vignettes/ggspatial.html).

## Spider Plots

The following code will generate a "spider plot" of all the samples that have a rock name of `"tuff"`, highlighting three samples. Modify the code below to give the plot:

- Appropriate `labs()` for the X, Y, and colour scales
- A log-Y axis
- Your favourite built-in ggplot `theme_*()`.

```{r}
warwick_meas <- warwick %>%
  filter(rock_type == "sedimentary") %>%
  gather(element, value, ends_with("ppm")) %>%
  mutate(element = element %>% str_remove("_ppm") %>% as_factor()) 

warwick_ref <- warwick_meas %>%
  filter(sample_id == "15GB0050") %>%
  select(element, reference_value = value)

warwick_norm <- warwick_meas %>%
  left_join(warwick_ref, by = "element") %>%
  mutate(reference = value / reference_value)

ggplot(warwick_norm) +
  geom_line(aes(x = element, y = reference, group = sample_id, color = rock_name), alpha = 0.3) +
  labs(x = NULL, y = "Concentration / Reference concentration", color = "Rock Name") +
  scale_y_log10() +
  theme_bw()
```

There are a number of new concepts in this code, including (1) using `gather()` to turn columns into rows (see the [tidying data tutorial](https://r4ds.had.co.nz/tidy-data.html)), (2) using `left_join()` to combine related datasets (see the [relational data tutorial](https://r4ds.had.co.nz/relational-data.html)), and (2) specifying a function as the `data` argument of a `geom_*()` function. Until you master all that, you should use this template, modifying the initial data (`warwick`) and the reference data (`warwick_ref`) to your needs.

## Stratigraphic Diagrams

The following code below creates a stratigraphic diagram for geochemical data from [Alta Lake, Whistler, British Columbia](https://doi.org/10.1007/s10933-016-9919-x) using the [tidypaleo package](https://github.com/paleolimbot/tidypaleo). This package makes the `geom_lineh()` and `facet_geochem_wraph()` functions available, as well as the `alta_lake_geochem` dataset. 

- Change the `geom_lineh()` to `geom_line()`. What is the difference between the two geom functions?
- Change the `facet_geochem_wraph()` function to `facet_wrap()`. What is the difference between the two facet functions?

```{r}
library(tidypaleo)

ggplot(alta_lake_geochem, aes(x = value, y = depth)) +
  geom_lineh() +
  geom_point() +
  theme_bw() + theme(strip.background = element_blank()) +
  facet_geochem_wraph(vars(param), nrow = 1, units = c("C" = "%", "Cu" = "ppm", "Ti" = "ppm")) +
  labs(y = "Depth (cm)", x = NULL) +
  scale_y_reverse()
```

If you are plotting species abundance data, the following example may be more relevant to you:

```{r}
keji_lakes_plottable %>%
  filter(location == "Beaverskin Lake") %>%
  ggplot(aes(x = rel_abund, y = depth)) +
  geom_col_segsh() +
  scale_y_reverse() +
  theme_bw() + theme(strip.background = element_blank()) +
  facet_abundanceh(vars(taxon)) +
  labs(y = "Depth (cm)")
```

## Stereonets

To my knowledge there isn't a good way to do this in ggplot2 and make the result pretty (yet!).

## Ternary Diagrams

Modify the code for this ternary diagram to 

- Map the `rock_type` variable to the `color` aesthetic. 
- Use the `labs()` function to modify the labels for K, Ca, and Ti.
- Apply your favourite built-in ggplot2 theme

```{r}
library(ggtern)

ggtern(warwick, aes(x = K_ppm, y = Ca_ppm, z = Ti_ppm, color = rock_type)) +
  geom_point(alpha = 0.2) +
  labs(x = "K", y = "Ca", z = "Ti", color = "Rock Type") +
  theme_bw()
```

For more on how to use the [ggtern package](http://ggtern.com), see the [journal article](https://www.jstatsoft.org/article/view/v087c03). Note that the ggtern package breaks other plots (including those from the `tidypaleo` package), so you might want to generate these plots separately! (you can choose _Restart R_ to restart your R session to a clean session uncontaminated by _ggtern_).
